Classless Java: Tuning Java Interfaces

### 2016-03-17 OOPSLA16/sections/Formalization.tex

\section{Interaction of Interface Methods with Interface Composition}
Before formalizing Classless Java and object interfaces, it is helpful
to informally discuss the behaviour of Java8
interfaces, in contrast with conventional trait models.

%We show several interesting cases
%and summarize compilation result of method composition into 3 categories:
%conflict error, accepted (both abstract) and conservative error. Meanwhile we
%also show different composition handling mechanism among traits, javac and ECJ.

\subsection{Methods in Java 8 Interfaces}
In Java 8 interfaces there are three types of methods: abstract, default, and static methods. Default and static methods were not allowed in interfaces in previous versions of Java.

\paragraph{Static methods} are handled in a very clean way: they are visible only in
  the interface in which they are explicitly defined. This means the following code
  is ill-typed.
\begin{lstlisting}
interface A0 { static int m(){return 1;} }
interface B0 extends A0 {}
... B0.m()//ill typed
\end{lstlisting}
This is different from the way static methods are handled in classes. Here
static methods have simply no interaction with interface
composition (\Q@extends@ or \Q@implements@).

\paragraph{Abstract method} composition is accepted when there exists a most specific one.
  For example, here method \texttt{Integer m()} from \texttt{B1} is visible in \Q@C1@.
\begin{lstlisting}
interface A1 { Object m(); }
interface B1 { Integer m();}
interface C1 extends A1,B1 {} //accepted
\end{lstlisting}

\paragraph{Default methods} conflict with any other default or abstract method. For
  example the following code is rejected due to method conflicts.
\begin{lstlisting}
interface A2 { default int m() {return 1;}}
interface B2 { int m(); }
interface C2 { default int m() {return 2;}}
interface D2 extends A2,B2 {} //rejected due to conflicting methods
interface E2 extends A2,C2 {} //rejected due to conflicting methods
\end{lstlisting}
Note how this is different from what happens in most trait models, where \Q@D2@
would be accepted, and the implementation in \Q@A2@ would be part of the
behavior of \Q@D2@.

\paragraph{Resolving conflicts:}
A method in the current interface wins over any method in its
super-interfaces, provided that the method
is the most specific one. This method also overrides conflict due to
inheritance. For example, the following code is accepted, but would be rejected
(see before) if the method \Q@m@ was not redefined in \texttt{D3} and
\texttt{E3}.
\begin{lstlisting}
interface D3 extends A2,B2 { int m(); } //accepted
interface E3 extends A2,C2 { default int m(){return 42;} } //accepted
\end{lstlisting}

\subsection{Classifying Outcomes of Interface Composition}
%When interfaces are composed and methods with the same name (and
%signature) exist, there are 3 possible outcomes.
%
We now try to classify possible outcomes for composition of methods with the same name (and
signature).
We will use the following (correct) declarations:
\begin{lstlisting}
interface A1 { T m(); }
interface A2 extends A1 { default T m(){ ... } }
interface A3 extends A2 { T m(); }

interface B1 { default T m(){ ... } }
interface B2 extends B1 { T m(); }
interface B3 extends B2 { default T m(){ ... } }
\end{lstlisting}

\noindent
What happens if a new interface \Q@M@ extends one \Q@A@${}_i$ and one
\Q@B@${}_j$?
% Nine representative cases are shown next:\\*
\begin{figure*}[htbp]
\centering
\begin{tabular}{|l|l|l|l|}
\hline
\textbf{M extends} & \textbf{A1}                  & \textbf{A2} & \textbf{A3} \\ \hline
\textbf{B1}        & conservative error                     & conflict error      & conservative error       \\ \hline
\textbf{B2}        & both abstract (accepted)     & conservative error       & both abstract (accepted)       \\ \hline
\textbf{B3}        & \textbf{conservative error} &conflict  error       & conservative error      \\ \hline
\end{tabular}
\end{figure*}
%We try to classify the results in the table:
\begin{itemize}
\item \textbf{conflict error} happens when the methods from both interfaces are implemented, which is also an error in most trait models.
\item \textbf{both abstract (accepted)} happens when the methods from both interfaces are abstract, which is also considered correct in all
  trait models.
\item \textbf{conservative error} happens when only one method is implemented
  (leaving another one abstract), which is different from what we would expect in
  a trait model, but is coherent with the conservative idea that a method
  defined in an interface should not silently satisfy a method in another one.
\end{itemize}

\paragraph{A bug:} During our experimentation, we found a bug in ECJ (Eclipse compiler for Java):
the case \textbf{M} extending \textbf{B3} and \textbf{A1} is accepted by
ECJ4.5.1 and rejected by javac.  By email communication with Brian Goetz
(leading Java 8 designer) we have confirmed that the expected behavior is
rejection, hence this is a bug in ECJ. This bug was also reported by
  others and is fixed in the ECJ developer branch, but not released as a stable
  version yet.
%Table~\ref{table:javabug} shows the method overriding bug in Java. In the
%example, there are 6 interfaces \texttt{A1,A2,A3,B1,B2,B3} with methods all
%named as \texttt{m} which are either abstract or default methods. If we define a
%new interface \texttt{M} that extends two of these interfaces, then the method
%overriding result is shown in the table. For example, row2 col2 means \texttt{M
%  extends A1,B1}. The result is \texttt{ERROR} because the abstract method
%\texttt{T m();} in \texttt{A1} conflicts with the method \texttt{default T m()
%  {...}} in \texttt{B1}. Readers may also figure out other extending cases by
%following this way of interpretation.

>>>>

%Table~\ref{table:javabug} shows the method overriding bug in Java. In the
%example, there are 6 interfaces \texttt{A1,A2,A3,B1,B2,B3} with methods all
%named as \texttt{m} which are either abstract or default methods. If we define a
%new interface \texttt{M} that extends two of these interfaces, then the method
%overriding result is shown in the table. For example, row2 col2 means \texttt{M
%  extends A1,B1}. The result is \texttt{ERROR} because the abstract method
%\texttt{T m();} in \texttt{A1} conflicts with the method \texttt{default T m()
%  {...}} in \texttt{B1}. Readers may also figure out other extending cases by
%following this way of interpretation.

### 2016-03-17 OOPSLA16/sections/Introduction.tex

Object oriented languages always strive to offer great code reuse,
they aim to couple flexibility and rigour, expressive power and modular reasoning.
Two main idea have emerged to this end: prototype based (PB) and class based (CB) languages.
In prototype based languages an objects inherits from other objects,
and thus objects own both behaviour and state (and objects is all you have).
In class based languages an object is instance of a specific class, and classes inherits from other classes.
Here objects own the state, while classes contains behaviour and the structure of the state.

@@

  State is just a special kind of behaviour, and is accessed by getter and setter methods.

  The literature provide good examples on how easy and modular can be combining
  multiple sources of pure behaviour (traits[]), and how
  hard it is to modularly combine multiple sources of behaviour \textbf{and} state (multiple inheritance[]).
  However, our object are the only responsible to define 
  the ultimate behaviour or a method, and if such method is just a getter/setter.

Our idea could be explained by defining a novel language, with new syntax and semantic.
However, this would have a
steep learning curve; for the sake of providing a more accessible explanation, we will encode
\footnote{
In the same sense, you can encode prototype based programming over class based with
the strategy pattern, and class based programming with prototypes by creating ``class'' objects.
} our ideas
over Java, relying on the new features of Java 8; namely
interface \emph{static methods}, and \emph{default methods}, allowing interfaces to have
method implementations. While the motivation for default methods was
to allow interfaces to be extended over time while preserving backwards
compatibility (\emph{interface evolution}),
 default methods can also emulate \emph{traits}~\cite{scharli03traits}.
\footnote{The
original notion of traits by Scharli et al. prescribes, among other
things, that: 1) a trait provides a set of methods that implement
behavior; and 2) a trait does not specify any state variables, so the
methods provided by traits do not access state variables
directly. Java 8 interfaces follow similar principles too. Indeed, a
detailed description of how to emulate trait-oriented programming in
Java 8 can be found in the work by Bono et al.~\cite{bono14}. The Java 8
team designing default methods, was also fully aware of that secondary
use of interfaces, but it was not their objective to model traits:
``The key goal of adding default methods to Java was "interface
evolution", not "poor man's traits"''~\cite{goetz13default}. As a result, 
they were happy to support the secondary use of interfaces with
default methods as long as it did not make the implementation and
language more complex.}

@@

We propose a programming style, where we never use classes\footnote{More precisely, we never use the \Q@class@ keyword.}.
We call this restricted version of Java \emph{Class-less Java}.
Our proposed style enables powerful object-oriented idioms, similar to 
multiple inheritance.

Since Java was not designed to be used in this way, our style can be verbose, especially about
object instantiation.
We will annotate with \Q|@Obj| certain interfaces and rely on
the Annotation processing to generate boring and repetitive code
allowing interface instantiation.
We call such annotated interface \emph{object interfaces};
they support \emph{abstract state operations}, providing a way
to use multiple inheritance with state in Java. The abstract state
operations include various common utility methods (such as getters and
setters, or clone-like methods). In the presence of subtyping, such
operations often require special care, as their types need to be
refined. Object interfaces provide support for type-refinement and can
automatically produce code that deals with type-refinement
adequately. %With object interfaces, many Java programs can be built
%without using a single class!

@@

@@

\item {\bf Object Interfaces:} A simple feature that allows various
  powerful multiple-inheritance programming idioms to be expressed
  conveniently in Java. We provide examples, informal and formal description.

>>>>

\bruno{need to defend ourselfves against Scala traits}
\marco{I tried to improve a little, also in this sense,
but I still think is better to defend ourselves mostly in the related work section, some of your text under may end up there}

\bruno{The abstract state operations and the novel technique to deal with 
covariant refinement of setters; Scala traits do support state (variables and 
immutable values), but there's no support for ``with'' (and clone-like) methods for example; and 
mutable state does not support type-refinement.
}

Object oriented languages always strive to offer great code reuse,
they aim to couple flexibility and rigour, expressive power and modular reasoning.
Two main idea have emerged to this end: prototype based (PB) and class based (CB) languages.
In prototype based languages an objects inherits from other objects,
and thus objects own both behaviour and state (and objects is all you have).

In class based languages an object is instance of a specific class, and classes inherits from other classes.
Here objects own the state, while classes contains behaviour and the structure of the state.

@@

  The literature provide good examples on how easy and modular can be combining
  multiple sources of pure behaviour (traits[]), and how
  hard it is to modularly combine multiple sources of behaviour \textbf{and} state (multiple inheritance[]).

  In IB, state is just a special kind of behaviour implementation.
  Object are the only responsible to define 
  the ultimate behaviour of a method and, for example, if such method is just a setter.
  Anything related to state is completely contained in the instances and do not leak in the interfaces.

Our idea could be explained by defining a novel language, with new syntax and semantic.
However, this would have a
steep learning curve; for the sake of providing a more accessible explanation, we will encode
\footnote{
In the same sense, you can encode prototype based programming over class based with
the strategy pattern, and class based programming with prototypes by creating ``class'' objects.
} our ideas
over Java, relying on the new features of Java 8; namely
interface \emph{static methods}, and \emph{default methods}, allowing interfaces to have
method implementations. While the motivation for default methods was
to allow interfaces to be extended over time while preserving backwards
compatibility (\emph{interface evolution}),
 default methods can also emulate \emph{traits}~\cite{scharli03traits}.
\footnote{The
original notion of traits by Scharli et al. prescribes, among other
things, that: 1) a trait provides a set of methods that implement
behavior; and 2) a trait does not specify any state variables, so the
methods provided by traits do not access state variables
directly. Java 8 interfaces follow similar principles too. Indeed, a
detailed description of how to emulate trait-oriented programming in
Java 8 can be found in the work by Bono et al.~\cite{bono14}. The Java 8
team designing default methods, was also fully aware of that secondary
use of interfaces, but it was not their objective to model traits:
``The key goal of adding default methods to Java was "interface
evolution", not "poor man's traits"''~\cite{goetz13default}. As a result, 
they were happy to support the secondary use of interfaces with
default methods as long as it did not make the implementation and
language more complex.}
\marcoT{We could use Scala Traits to encode IB, but the simpler Java8 model is a better starting point.}
\footnote{
\marco{may go in related work}
Scala traits can declare \Q@def, val, var@ and \Q@lazy val@ 
However, while \Q@def@s can be overridden by \Q@val@/\Q@var@s, 
\Q@val@/\Q@var@s can not be overridden.\marco{we need to check this}
A\Q@val@/\Q@var@ member is predestined to be instance state, and this
is in sharp contrast with our model where the state is only a private detail of instances, and
do not contaminate interface/trait/classes.}
\footnote{Scala traits have no constructor, but 
\Q@var@s and (\Q@lazy@) \Q@val@s can be ``pre-initialized''.}

@@

We propose a programming style, where we never use classes\footnote{More precisely, we never use the \Q@class@ keyword.}.
We call this restricted version of Java \emph{Class-less Java}.
Our proposed style enables powerful object-oriented idioms, similar to 
(stateful) multiple inheritance and field type refinement 
(allowing a kind of covariant setters refinement).

As for CB, in IB programmers do not define objects directly,but delegate the task to
\emph{object interfaces}, whose role is similar to non abstract classes in CB.

The idea is that all abstract operations in an Object interface will be
interpreted as \emph{abstract state operations};  The abstract state
operations include various common utility methods (such as getters and
setters, or clone-like methods). In the presence of subtyping, such
operations often require special care, as their types need to be
refined. Object interfaces provide support for type-refinement and can
automatically produce code that deals with type-refinement
adequately, while Java (and Scala) require explicit type-refinement.

  %With object interfaces, many Java programs can be built
%without using a single class!

The Java annotation processors allows us to stick to pure Java8: by
annotating with \Q|@Obj| certain interfaces
we can generate boring and repetitive code
allowing interface instantiation and type refinement; such code should not be needed in the first place
in a real IB language.
%Since Java was not designed to be used in this way, our style can be verbose, especially about object instantiation.

@@

We can extend our approach to support encapsulation and privateness;
various possibilities are discussed in the last part of the paper.

@@

\item {\bf IB and Object Interfaces:} a novel take on object orientation, allowing
  powerful multiple-inheritance programming idioms to be expressed
  conveniently in Java. We provide examples, informal and formal description.

### 2016-03-17 OOPSLA16/sections/Overview.tex

This section illustrates how our programming style, supported by the \mixinAnn{} annotation, enable powerful
idioms similar to multiple inheritance and field type refinement.
We propose a standard example: \Q@Animal@s with a two-dimensional \Q@Point2D@ representing their \Q@location@.
Some kinds of animals are \Q@Horses@ and \Q@Bird@s.
Birds can \Q@fly@, thus their locations need to be three-dimensional \Q@Point3D@s (field type refinement).
Finally, we model \Q@Pegasus@ (one of the best-known creatures in Greek
mythology) as a kind of \Q@Animal@ with the skills
of both \Q@Horse@s and \Q@Bird@s (multiple inheritance). A simple class diagram illustrating
the basic system is given on the left side of Figure~\ref{fig:pegasus}.\footnote{Some research argues in favor of using subtyping for modeling taxonomies, other research argues against this practice, we do not wish to take sides in this argument, but to provide an engaging example.}
\bruno{Should we provide a one sentence summary in the abstract of how much code
is needed in Java (without Obj) vs the approach with CJ, for the Pegasus example?}

\subsection{Simple Multiple Inheritance with Default
  Methods}\label{sec:simple}

@@

\begin{lstlisting}
interface Point2D { int x(); int y();
    static Point2D of(int _x, int _y){ return new Point2D(){
        int x=_x; int y=_y;
        public int x(){return x;}   public int y(){ return y; }
        Point2D withX(int val){ return Point2D.of(val,this.y()); }
        Point2D withY(int val){ return Point2D.of(this.x(),val); }
    }; }
  Point2D withX(int val);    Point2D withY(int val); }
\end{lstlisting}
\end{comment}
Moreover, if the programmer wants a different implementation, he may
provide an alternative implementation using \Q@default@ methods. For example:
\begin{lstlisting}
@Obj interface Point2D {
    int x(); int y();
    default Point2D withX(int val){ ... }
    default Point2D withY(int val){ ... } }
\end{lstlisting}

%\begin{comment}
%\marco{I re-enabled this code, I think is needed for understandability}
\noindent is expanded into
\begin{lstlisting}
interface Point2D {
    int x(); int y();
    default Point2D withX(int val){ ... }
    default Point2D withY(int val){ ... }
    static Point2D of(int _x, int _y){
      return new Point2D(){
        int x=_x;    int y=_y;
        public int x(){return x;}
        public int y(){return y;} }; } }
\end{lstlisting}

@@

\paragraph{Summary} Dealing with state (mutable or not) in object interfaces
relies on a notion of abstract state, where only \emph{methods} that interact with
state are available to users. Object interfaces provide support for four
different types of abstract state operations \textcolor{red}{together with two types of
general operations}, which are summarized in
Figure~\ref{fig:abstractstate}.  The abstract state operations are determined by
naming conventions and the types of the methods. Fluent setters are a variant of
conventional setters, and are discussed in more detail in Section~\ref{sec:dsls}.

\begin{figure*}
\centering
\saveSpaceFig
\begin{tabular}{|l|l|l|l|}
\hline
& \textbf{Operation}  & \textbf{Example}                  & \textbf{Description } \\ \hline
\multirow{4}{*}{State operations} & \textbf{``fields''/getters}        &   \Q@int x();@                  & Retrieves value from field \texttt{x}.          \\ \cline{2-4}
& {\bf withers}        &   \Q@Point2D withX(int val);@                & Clones
object; updates field \texttt{x} to \texttt{val}.             \\ \cline{2-4}
& \textbf{setters}        & \Q@void x(int val);@ & Sets the field
\texttt{x} to a  new value \texttt{val}.        \\ \cline{2-4}
& \textbf{fluent setters}        & \Q@Point2D x(int val);@ &Sets the field
\texttt{x} to \texttt{val} and returns \texttt{\this}.           \\ \hline
\multirow{2}{*}{Other operations} & \textbf{factory methods} & \begin{tabular}{l}\Q@static Point2D of(int _x,@\\ \hspace{.8in}\Q@int _y) {...}@\end{tabular} & Generate factory method for \texttt{Point2D}.        \\ \cline{2-4}
& \textbf{general updaters}        & \Q@Point3D with(Point2D val);@ & Update all fields
that match the object \texttt{val}.        \\ \hline
\end{tabular}

\caption{\textcolor{red}{(edited)} Abstract state operations for a field \texttt{x}, together with general operations, supported by the \mixin
  annotation. \marco{I disagree with this table, for example we miss the with} }

@@

\noindent From the type perspective, the key is the covariant method
overriding of \texttt{location()}. However, from the semantics
perspective the key is the implementation for the setter with the old
signature (\Q@location(Point2D)@). The key to the setter
implementation is a new type of \Q@with@ method, called a property updater.

\bruno{At some point we should talk about the point that was discussed in the reviews}
\paragraph{\Q@Point3D@ and properties updaters}
The \Q@Point3D@ interface is defined as follows:

@@

\begin{figure}
\saveSpaceFig
\lstinputlisting[linerange=79-91]{../UseMixinLombok/src/pegasus/TestAnimal.java}% APPLY:linerange=GENERATED_POINT3D
\caption{Generated boilerplate code.}
\label{fig:boilerplate}
\end{figure}

\paragraph{Generated boilerplate}
Just to give a feeling of how much mechanical code \mixin is generating, we show the
generated code for the \texttt{Point3D} in Figure~\ref{fig:boilerplate}. Writing such code by hand is error-prone. For
example a distracted programmer may swap the arguments of calls to
\Q@Point3D.of@.  Note how \Q@with-@ methods are automatically refined in their
return type, so that code like:

>>>>

This section illustrates how our programming style, supported by the \mixinAnn{} annotation, enable powerful
idioms similar to multiple inheritance and field type refinement.
We propose a standard example: \Q@Animal@s with a two-dimensional \Q@Point2D@ representing their \Q@location@.
Some kinds of animals are \Q@Horses@ and \Q@Bird@s.
Birds can \Q@fly@, thus their locations need to be three-dimensional \Q@Point3D@s (field type refinement).
Finally, we model \Q@Pegasus@ (one of the best-known creatures in Greek
mythology) as a kind of \Q@Animal@ with the skills
of both \Q@Horse@s and \Q@Bird@s (multiple inheritance). A simple class diagram illustrating
the basic system is given on the left side of Figure~\ref{fig:pegasus}.\footnote{Some research argues in favor of using subtyping for modeling taxonomies, other research argues against this practice, we do not wish to take sides in this argument, but to provide an engaging example.}
\bruno{Should we provide a one sentence summary in the abstract of how much code
is needed in Java (without Obj) vs the approach with CJ, for the Pegasus example?}
\marco{Not in the abstract. But I think is good to have that number and to put in in the example section.
I will set up the sentence and someone can compute the number.}
\subsection{Simple Multiple Inheritance with Default
  Methods}\label{sec:simple}

@@

\begin{lstlisting}
interface Point2D { int x(); int y();
    static Point2D of(int _x, int _y){ return new Point2D(){
        int x=_x; int y=_y;
        public int x(){return x;}   public int y(){ return y; }
        Point2D withX(int val){ return Point2D.of(val,this.y()); }
        Point2D withY(int val){ return Point2D.of(this.x(),val); }
    }; }
  Point2D withX(int val);    Point2D withY(int val); }
\end{lstlisting}
\end{comment}
Moreover, if the programmer wants a different implementation, he may
provide an alternative implementation using \Q@default@ methods. For example:
\begin{lstlisting}
@Obj interface Point2D {
    int x(); int y();
    default Point2D withX(int val){ /*myCode*/ }
    default Point2D withY(int val){ /*myCode*/ } }
\end{lstlisting}

%\begin{comment}
%\marco{I re-enabled this code, I think is needed for understandability}
\noindent is expanded into
\begin{lstlisting}
interface Point2D {
    int x(); int y();
    default Point2D withX(int val){ /*myCode*/ }
    default Point2D withY(int val){ /*myCode*/ }
    static Point2D of(int _x, int _y){
      return new Point2D(){
        int x=_x;    int y=_y;
        public int x(){return x;}
        public int y(){return y;} }; } }
\end{lstlisting}

@@

\paragraph{Summary} Dealing with state (mutable or not) in object interfaces
relies on a notion of abstract state:
\marcoT{Method bodies can only refer to other methods;
Objects are (conceptually) closures and the state is 
composed by captured local variables.}
Methods interacting with those are called \emph{abstract state operations}.

Object interfaces provide support for four
different types of abstract state operations \textcolor{red}{together with two types of
general operations}, which are summarized in
Figure~\ref{fig:abstractstate}.  The abstract state operations are determined by
naming conventions and the types of the methods. Fluent setters are a variant of
conventional setters, and are discussed in more detail in Section~\ref{sec:dsls}.

\begin{figure*}
\centering
\saveSpaceFig
\begin{tabular}{|l|l|l|l|}
\hline
& \textbf{Operation}  & \textbf{Example}                  & \textbf{Description } \\ \hline
\multirow{4}{*}{State operations} & \textbf{``fields''/getters}        &   \Q@int x()@                  & Retrieves value from field \texttt{x}.          \\ \cline{2-4}
& {\bf withers}        &   \Q@Point2D withX(int val)@                & Clones
object; updates field \texttt{x} to \texttt{val}.             \\ \cline{2-4}
& \textbf{setters}        & \Q@void x(int val)@ & Sets the field
\texttt{x} to a  new value \texttt{val}.        \\ \cline{2-4}
& \textbf{fluent setters}        & \Q@Point2D x(int val)@ &Sets the field
\texttt{x} to \texttt{val} and returns \texttt{\this}.           \\ \hline
\multirow{3}{*}{Other operations} &
\textbf{factory methods} & 
%\begin{tabular}{l}
\Q@static Point2D of(int _x,int _y)@
%\end{tabular}
 & Generate factory method for \texttt{Point2D}.        \\
\cline{2-4}
& \textbf{functional updaters}        & \Q@Point3D with(Point2D val)@ & Update all fields
that match fields in \texttt{val}.        \\
\cline{2-4}
& \textbf{imperative updaters}        & \Q@Point3D set(Point2D val)@ & Set all fields
that match fields in \texttt{val}.        \\ \hline
\end{tabular}

\caption{\textcolor{red}{(edited2)} Abstract state operations for a field \texttt{x}, together with general operations, supported by the \mixin
  annotation. }

@@

\noindent From the type perspective, the key is the covariant method
overriding of \texttt{location()}. However, from the semantics
perspective the key is the implementation for the setter with the old
signature (\Q@location(Point2D)@). The key to the setter
implementation is a new type of \Q@with@ method, called
 a (functional) property updater.

\bruno{At some point we should talk about the point that was discussed in the reviews}
\marco{you mean the reviewer question on the list.size?}
\paragraph{\Q@Point3D@ and properties updaters}
The \Q@Point3D@ interface is defined as follows:

@@

 \marcoT{Symmetrically,
 we could offer an imperative property updater
 that calls the setters instead of the withers.
 \Q@Point3D set(Point2D val)@.
 }

 \marcoT{
 Property updaters never break class invariants, since they
 internally call operations that was already deemed
 safe by the programmer. For example a list object
 would not offer a setter for its size field, thus
 a property updater would not attempt to set it.}

\begin{figure}
\marco{I can not find the declaration int x();
also we can change the order of the code to write the non generated stuff first.}
\saveSpaceFig
\lstinputlisting[linerange=79-91]{../UseMixinLombok/src/pegasus/TestAnimal.java}% APPLY:linerange=GENERATED_POINT3D
\caption{Generated boilerplate code.}
\label{fig:boilerplate}
\end{figure}

\paragraph{Generated boilerplate}
Just to give a feeling of how much mechanical code \mixin is generating, we show the
generated code for the \texttt{Point3D} in Figure~\ref{fig:boilerplate}.
\marcoT{Overall, for the whole 
animals-with-locations, an @@@ lines code example,
the generate/completed code is composed by
@@@ lines.}

   Writing such code by hand is error-prone. For
example a distracted programmer may swap the arguments of calls to
\Q@Point3D.of@.  Note how \Q@with-@ methods are automatically refined in their
return type, so that code like:
